---
title: "BioResonator Subgroup: Using R for Data Analysis"
author: "James Wade"
date: "November 11, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Using R to Plot Compiled Data

Here is the the code to analyze compiled data from multiplex protein detection in patient derived tumor xenografts under various treatments. 

Here is the code to access the data from GitHub:

```{r, echo = TRUE}
url <- "https://raw.githubusercontent.com/jwade1221/
XenograftProteinProfiling/master/Combined_Averaged.csv"
filename <- basename(url)
download.file(url,destfile=filename)
```

## Get the Data

Importing the data using `library(readr)` and `read_csv` (plus a look at the data file):

```{r, echo = TRUE}
library(readr)
dat <- read_csv(filename, col_types = cols())
head(dat)
```

## Function for Plotting Each Target

```{r, echo = TRUE, eval = FALSE}
plotEachTarget <- function(i){
    library(readr)
    library(ggplot2)
    targetList <- unique(dat$Target)
    dat.Target <- dat[dat$Target == targetList[i],]
    
    #set error bars
    limits <- aes(ymax = `Relative Shift` + `Standard Deviation`, 
                  ymin = `Relative Shift` - `Standard Deviation`)
```

## Function for Plotting Each Target
```{r, echo = TRUE, eval= FALSE}
#plot figure
plots <- ggplot(data = dat.Target,
    aes(x = Treatment, y = `Relative Shift`, 
        group = interaction(factor(`Time Point`), 
        factor(`Cell Line`)))) +
    geom_bar(stat = "identity", 
        aes(fill = interaction(factor(`Time Point`), 
        factor(`Cell Line`))), 
        position = position_dodge(width = 0.9)) +
    ylab("Net Shift (pm)") +
    geom_errorbar(limits, width = 0.4, size = 0.3,
        position = position_dodge(width = 0.9)) +
    theme_bw() + theme(panel.grid = element_blank(), 
        axis.title.x=element_blank()) + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle(paste("Target: ", targetList[i], sep=''))
```

## Function for Plotting Each Target
   
```{r, echo = TRUE, eval = FALSE}
#configure legend
plots <- plots + 
        scale_fill_discrete(name="Cell Line and Treatment", 
        labels=c("GBM-6 1 h Treatment", "GBM-6 24 h Treatment", 
        "GBM-26 1 h Treatment", "GBM-26 24 h Treatment")) + 
        theme(legend.key = element_rect(colour = 'white',
        fill = 'white'), legend.key.size = unit(0.4, "cm")) 
#plot figure
plots

#save figure
filename = paste(targetList[i], 'png', sep='.')
ggsave(plots, file = filename, width = 8, height = 6)
}

#execute function
plotEachTarget(3)
```

## Plotting by Target

```{r}
plotEachTarget <- function(i){
    library(readr)
    library(ggplot2)
    targetList <- unique(dat$Target)
    dat.Target <- dat[dat$Target == targetList[i],]
    
    #set error bars
    limits <- aes(ymax = `Relative Shift` + `Standard Deviation`, 
                  ymin = `Relative Shift` - `Standard Deviation`)
    
    #plot figure
    plots <- ggplot(data = dat.Target,
        # group lets `ggplot` know we want different errorbars/bars for each day
        aes(x = Treatment, y = `Relative Shift`, 
            group = interaction(factor(`Time Point`), factor(`Cell Line`)))) +
        geom_bar(stat = "identity", aes(fill = interaction(factor(`Time Point`), 
            factor(`Cell Line`))), position = position_dodge(width = 0.9)) +
        ylab("Net Shift (pm)") +
        geom_errorbar(limits, width = 0.4, size = 0.3,
            position = position_dodge(width = 0.9)) +
        theme_bw() + theme(panel.grid = element_blank(), 
            axis.title.x=element_blank()) + 
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        ggtitle(paste("Target: ", targetList[i], sep=''))

    #configure colors and legend
    cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", 
                   "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
    plots <- plots + scale_fill_manual(name="Cell Line and Treatment", 
                labels=c("GBM-6 1 h Treatment", "GBM-6 24 h Treatment", 
                "GBM-26 1 h Treatment", "GBM-26 24 h Treatment"), values = cbPalette) + 
                theme(legend.key = element_rect(colour = 'white',
                    fill = 'white'), legend.key.size = unit(0.4, "cm"))  
    plots            
    #save figure
    #filename = paste(targetList[i], 'png', sep='.')
    #ggsave(plots, file = filename, width = 8, height = 6)
}
plotEachTarget(3)
```

## Plotting by Time Point & Cell Line

```{r}
plotEachTimepoint <- function(i, j){
    
    #create lists for data islation
    timepointList <- unique(dat$`Time Point`)
    cellLineList <- unique(dat$`Cell Line`)
    dat.Timepoint <- dat[dat$`Time Point` == timepointList[i] & 
                             dat$`Cell Line` == cellLineList[j],]
    
    #set error bars
    limits <- aes(ymax = `Relative Shift` + `Standard Deviation`, 
                  ymin = `Relative Shift` - `Standard Deviation`)
    
    #plot colors
    extPalette <- c("#771155", "#AA4488", "#CC99BB", "#114477", "#4477AA", 
                   "#77AADD", "#117777", "#44AAAA", "#77CCCC", "#117744", 
                   "#44AA77", "#88CCAA", "#774411", "#AA7744", "#DDAA77", 
                   "#771122", "#AA4455", "#DD7788")
    
    #plot figure settings
    plots <- ggplot(data = dat.Timepoint,
        # group lets `ggplot` know we want different errorbars/bars for each day
        aes(x = Treatment, y = `Relative Shift`, group = factor(Target))) +
        geom_bar(stat = "identity", position = position_dodge(width = 0.9), 
                 aes(fill = Target)) +
        ylab("Net Shift (pm)") +
        geom_errorbar(limits, width = 0.4, size = 0.3, colour = 'grey48',
                      position = position_dodge(width = 0.9)) +
        theme_bw() + theme(panel.grid = element_blank(), 
                           axis.title.x=element_blank()) + 
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        ggtitle(paste(timepointList[i], " Hour Treatment ", "GBM-", cellLineList[j], sep=''))
    
    #format plot legend and colors
    plots <- plots + theme(legend.key = element_rect(colour = 'white', 
                fill = 'white'), legend.key.size = unit(0.4, "cm")) +
                scale_fill_manual(values = extPalette)
    plots
    #save figure
    #filename = paste(timepointList[i],cellLineList[j], 'png', sep='.')
    #ggsave(plots, file = filename, width = 8, height = 6)
}
plotEachTimepoint(1,1)
```


